---
title: "test_3_Andrew_Martinez"
author: "Andrew Martinez"
date: "12/20/2018"
runtime: shiny
output: html_document
---

## Install Necessary Packages
```{r install, include=FALSE}
if(!"quantmod" %in% installed.packages()) {install.packages("quantmod")}
library("quantmod")

if(!"readxl" %in% installed.packages()) {install.packages("readxl")}
library(readxl)

if(!"plotly" %in% installed.packages()) {install.packages("plotly")}
library("plotly")

if(!"reshape" %in% installed.packages()) {install.packages("reshape")}
library(reshape)

if(!"cluster" %in% installed.packages()) {install.packages("cluster")}
library(cluster) 

if(!"fpc" %in% installed.packages()) {install.packages("fpc")}
library(fpc)

if(!"NbClust" %in% installed.packages()) {install.packages("NbClust")}
library("NbClust")

if(!"magrittr" %in% installed.packages()) {install.packages("magrittr")}
library(magrittr)

if(!"data.table" %in% installed.packages()) {install.packages("data.table")}
library(data.table) 

if(!"shiny" %in% installed.packages()) {install.packages("shiny")}
library(shiny)

```


## Helper Functions
```{r function, include = FALSE}
#Change column labels for daily return and volatility to include associated ticker
includeTicker<-function(env_input, symbol){
  all_labels<-names(env_input[[symbol]])
  all_labels[length(all_labels)]<-paste0(symbol,".Daily_Return")
  
  #Override existing names with newly created label
  names(env_input[[symbol]]) <- all_labels
  output<-env_input[[symbol]]
  
  return(output)
  
}

#Plots charts related to regression residuals
residualAnalysis<-function(data_set, regression_output){
  
  #Histogram of Residuals
  hist(regression_output$residuals, main="Residual Histogram")
  
  #Boxplot of Residuals
  boxplot(regression_output$residuals, main="Residual Boxplot")
  
  #Line plot of Residuals
  plot(regression_output$residuals, type='l', main='Model Residuals')
}

#Load equity pricing data from yahoo from the start of the year 2016 to the present
getEquityData<-function(symbol, start){
  #Gives current status of data load
  cat("Downloading time series for symbol '", symbol, "' ...\n", sep = "")
  
  #Get data from yahoo and saves it in the NASDAQ environment
  getSymbols(symbol, env = nasdaq, src = "yahoo", from = as.Date(start))
  
  #Calculate the daily return
  nasdaq[[symbol]]$Daily_Return<-dailyReturn(nasdaq[[symbol]])
  
  #Rename newly created columns to include ticker name and extract only those columns into an output xts object
  final_output<-includeTicker(nasdaq, symbol)
  final_output<-final_output[,grep("Daily_Return",colnames(final_output))]

  return(final_output)
  
}
```


## Load Data
Loads all data from a provided list using the previously described helper functions.
```{r data_load, echo=FALSE, warning = FALSE}
#Prevent standard yahoo warnings from appearing
options("getSymbols.yahoo.warning"=FALSE)
options("getSymbols.warning4.0"=FALSE)

#Load ticker mapping from .xlsx file (Set WD as needed for nasdaq file)
ticker_mapping <- data.frame(read_xlsx("nasdaq_symbols.xlsx"))

#Set Start Date
starting_date = '2016-01-01'

#Create NASDAQ environment to make full use of quantmod functionality and to store original data
nasdaq <- new.env()
stocks <- lapply(ticker_mapping$Symbol, getEquityData, starting_date)
  
```


## Prep data for analysis
```{r cluster_prep, message = FALSE}
#Merges list of equity data into single xts data structure
only_returns<-do.call(merge.xts, stocks)

#Gets the standard deviation of the daily returns of each ticker across the full time horizon, converts the xts to a dataframe and reshapes the structure
sd_values<-melt(data.frame(lapply(only_returns, sd, na.rm=T)))
sd_values<-sd_values[,c("value")]

#Gets the mean of the daily returns of each ticker across the full time horizon, converts the xts to a dataframe and binds this structure the previously created structure containing standard deviation.
data_cluster<-cbind(melt(data.frame(lapply(only_returns, colMeans, na.rm=T))), sd_values)

#Renames the column headers and removes the ticker header
all_titles<-c("Ticker", "Avg_Daily_Return", "Volatility")
names(data_cluster)<-all_titles

#Sets the ticker as the row names and deletes the ticker column
row.names(data_cluster)<-ticker_mapping$Symbol
data_cluster$Ticker = NULL
head(data_cluster[order(data_cluster$Avg_Daily_Return, decreasing = T),])
```

## Regression Analysis
Regression analysis reveals a very weak relationship between the average daily return and volatility across the given the time horizon. Residuals follow a normal distribution, but display levels of heteroskedasticity. This is further demonstrated in graphical analysis of the line of best fit against the data. 
```{r regression}
#Create separate regression data structure to allow for custom manipulation
regression_analysis<-data.table(data_cluster)

#Run regression model and print summary
lm1 <- lm(Avg_Daily_Return ~ Volatility, data=regression_analysis)
print(summary(lm1))

#Adds output from regression model to the dataframe
df_fit<-regression_analysis[, lm_fit:=fitted(lm1)]

#Run Residual Analysis Function
residualAnalysis(df_fit, lm1)

#Scatter plot fo data with line of best fit
ggplot(df_fit, aes(x = Volatility, y = Avg_Daily_Return)) + geom_point() + ggtitle("Volatility vs Average Daily Return") + 
      stat_smooth(method = "lm", formula = y ~ x) 

```


## Find Optimal Cluster Count
```{r cluster count}
set.seed(62592)

#Using all 30 different indicies, determine the optimal number of clusters using euclidean distance (set range of possible clusters as 2 to 10). Values are scaled for normalization.
res.nbclust <- data_cluster %>%
  scale() %>%
  NbClust(distance = "euclidean",
          min.nc = 2, max.nc = 10, 
          method = "kmeans", index ="all") 

```


## K-means
Although the previous analysis largely recommended 3 clusters as the optimal set, 4 were ultimately used as the three clusters that were formed were of limited value (i.e. essentially split into a more aggressive, conservative and balanced clusters with regards to standard deviation with a mix of average returns). Using a 4 cluster scheme, one particular cluster that minimized the standard deviation while still having decent returns (cluster 1) contains this analysis's recommnended stocks for a balanced strategy. Cluster 3 represents a more aggressive strategy, with cluster 4 demonstrating relative underperformance and cluster 2 being a more conservative strategy. A clustering plot using PCA was shown as well, however it is difficult to intuitively understand its output. 
```{r kmeans, warning = FALSE}

#Runs kmeans algorithm with 4 clusters (only takes numeric variables and scales them)
k=4
fit<-kmeans(scale(data_cluster[, sapply(data_cluster, is.numeric)]), centers = k)
print(fit)

#Creates new scaled dataframe that contains the appended cluster mapping to original input data frame
equity.scaled <- data_cluster[, sapply(data_cluster, is.numeric)]
equity.scaled$clusters<-as.factor(fit$cluster)

colorsmap <- c('red','blue','green','orange')

#Regular Cluster Plot Scaled
plot_ly(equity.scaled, type = "scatter", x = ~Avg_Daily_Return, y = ~Volatility,
             text =row.names(equity.scaled),color = ~clusters, 
             colors = colorsmap, mode = 'markers') %>%
  
        layout(title = 'Equity Average Daily Return vs Volatility',
               yaxis = list(zeroline = FALSE),
               xaxis = list(zeroline = FALSE)) 

```


## Show plots of stocks of interest
```{r cluster_plots, echo = FALSE}
#Gets tickers for the stocks that were part of cluster 5 (balanced strategy)
balanced_list<-rownames(equity.scaled[equity.scaled$clusters == 1,])

#Show charts for the balanced list stocks using default quantmod chart settings in a Shiny App
shinyApp(
  ui = fluidPage(titlePanel("Stock Closing Price"),
  #Although analysis only goes as far back as 2016, balanced list plots allow for charting starting at 2000.
         sidebarLayout(sidebarPanel(
         dateInput(inputId = "start_date",label = "Start date",
                     min = '2000-01-01',
                     max = Sys.Date(),
                     value = starting_date),
         dateInput("end_date",
                   "End Date",
                   min = starting_date,
                   max = Sys.Date(),
                   value = Sys.Date()),
         selectInput("symbol", label = "Ticker", choices = balanced_list)),
      
      mainPanel(plotOutput("stockPlot")))),

  server = function(input, output) {
    #Recreate nasdaq environemnt to allow use of charSeries function
     nasdaq2 <- new.env()
     output$stockPlot <- renderPlot({
     stock_data<-quantmod::getSymbols(input$symbol, env = nasdaq2, src = "yahoo", from=input$start_date, 
     to=input$end_date)
     chartSeries(nasdaq2[[input$symbol]], name = paste0(input$symbol), TA="addVo();addMACD();addBBands()")
     
   })
  },

  options = list(height = 1000)
)
```
